---
title: "Tidy DISA Viral Load Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tidy-disa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Open RStudio and Load Packages

If not already done, open a new RStudio work session.  As normal we'll begin our work by creating a new R script (File -> New File -> R script).  We'll also load the core R packages needed for tidying DISA tabular data. Tidyverse is our trusty multipurpose "Swiss Army Knife" of data science programming in R.  The sismar package allows us to tidy data exported from MISAU health information systems including DISA. The glue package allows us to combine and use string values within code.


```{r, eval = FALSE, echo = TRUE}
library(tidyverse)
library(sismar)
library(glue)
```

## Define Global Variables

Before we process our DISA viral load data, let's set a couple values in the global R environment so that we can later pass them into our sismar function.  First define a value for the path to the OpenLDR DISA viral load .xlsx export saved on your local machine.  Define a second value for the month of the DISA viral load export from the OpenLDR system.  

Note that the month date needs to be set in format %Y-%m-%d ("YYYY-MM-DD"). Although there are no restrictions on defining the exact day of the date string it is best to be consistent if tidy'ing and compiling submissions over multiple months.  For example, if we're defining day as "01" in our processing of May DISA data then we should also define it as "01" in our June processing.  


```{r}
path_disa_cv <- "Data/Relatorio Mensal de Carga Viral Junho 2024.xlsx"
month_input <- "2024-06-20"

path_disa_cv
month_input

```

## Tidy Monthly DISA Viral Load Dataset

Can you believe we've already done the hard work of processing our monthly DISA viral load data set?!  All we need to do now is write a single line of code to call the sismar `process_disa_cv` function with the arguments we've already set in our global environment and we've got our tidy viral load data ready for analysis!

```{r, eval = FALSE}
df <- process_disa_cv(path_disa_cv)
```

## Write Monthly Dataset to Disk

You'll likely want to write your processed monthly dataset to disk.  This will be useful if (rather "when") you decide to analyze multiple months of DISA data together.  We can save .csv files to disk easily using `write_csv` from the Tidyverse readr package.  All we do is call `write_csv` and pass in our "df" data object along with the file connection information (location and name of the .csv we save to disk).

It's worth noting in the below code chunk that we've also managed to apply a handy trick!  We've used the glue package to append a period suffix automatically to the name of the .csv we save!  This will ensure that every month we do not overwrite previous months files and that we end up with a full set of uniquely named monthly files to compile together! Clever how we can embed functions within other functions to make both more powerful, right?!


```{r, eval = FALSE}
write_csv(df, file = glue("Data/processed/disa_vl_{month_input}")
```

## Compile Monthly Datasets

To append our saved monthly files together we're going to up the level of difficulty by using an R regular expression and by calling a Tidyverse purrr function called "map" that helps us iterate tasks (in this case the task of reading a .csv file).  At the end of the code chunk, we'll use the function "reduce" to append our .csv's together.

The most opaque code line in the chunk begins with, "list.files". This is where we effectively ask R to take an inventory of the complete file names in the folder where we'd saved our monthly processed DISA data.  The "pattern" argument allows us to put parameters around what we inventory.  The "^" symbol followed by "disa_vl_" ensures that only files that start with "disa_vl_" will be included in our compiled dataset.  

```{r, eval = FALSE}
df_compiled <-
  list.files("Data/processed/", pattern = "^disa_vl_", full.names = TRUE) %>%
  map(~ read_csv(.x)) %>%
  reduce(rbind)
```

## Celebrate (and analyze) Your Work!

We've done some nice work here so take a moment to give yourself a pat on the back!  However, don't pat for too long because now the really fun job begins of exploring the dataset and generating insights!  Happy analysis!
