---
title: "Tidy SISMA tabular program data"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{tidy_sop}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>"
)
```

## Open RStudio and Load Packages

Let's begin by opening a new RStudio work session.  Once a session is open, create a new R script (File -> New File -> R script) and load the core R packages needed for tidying SISMA .csv exports. The Tidyverse is a collection of open-source packages designed for general data science use. Sismar is a package designed for the specific purpose of munging and conducting feature engineering for SISMA tabular exports. Finally, the Glue packages provides functions that allow us to combine and use string values within code.

```{r, eval = FALSE, echo = TRUE}
library(tidyverse)
library(sismar)
library(glue)
```

## Define Global Variables

We'll begin our adventure by coding three values into our global environment for later use - 1) the calendar year for which data was downloaded from SISMA; 2) the path to the SISMA .csv export saved on our local machine; and 3) the path and filename where our processed dataset will be saved. Note that best practice is to store data and scripts in specific pre-created R project folder structures. This facilitates reproducibility and collaboration across users.   

```{r}
year <- "2024"
path_hiv_tarv <- glue::glue("Data/tarv_{year}.csv")
output_hiv_tarv <- glue::glue("Dataout/hiv_tarv_{year}.txt")

path_hiv_tarv
output_hiv_tarv
```

## Tidy SISMA Dataset

Having defined the local path to our SISMA dataset, we can now call functions from the sismar package to generate a tidy analytical dataframe.  This can be done either by using a combination of sismar functions called in sequence or by calling a singular wrapper function and specifying the appropriate arguments.  Let's start with the first, slightly more verbose, approach.  

##### Approach 1: Using Sequential Functions

We'll start by calling `clean_sisma_csv` and providing the function argument that specifies where our downloaded SISMA .csv export is saved on our local machine. Note that this function is not dependent on the type of program data contained within the .csv file. This means that clean_sisma_csv` can be used with any SISMA tabular download. In the output notice that the data frame's indicators have been pivoted to a long format, that a small number of variables have been removed, and that the names of the pivoted indicators have been modified to be more machine readable.

```{r, eval = FALSE}
df <- clean_sisma_csv(path_hiv_tarv)
```

Next, we'll call a sismar function to engineer data features that are encoded in the indicator names created by `clean_sisma_csv`.  In this use case, we call `parse_sisma_hiv_tarv`. It's important to note that there are numerous types of standard programmatic data sets that can be generated through saved SISMA reports. This means that the specific sismar parsing function called for feature engineering will vary depending on type of programmatic data set provided by the user. A full list and description of each sismar parsing function is provided in the sismar [package documentation](https://usaid-mozambique.github.io/sismar/reference/index.html).

```{r, eval = FALSE}
df <- parse_sisma_hiv_tarv(df)
```
In the parsing function output notice that new variables have been added to the data frame and that indicator names have been encoded to be more succinct. This tidy format is perfect for analysis!

##### Approach 2: Using Wrapper Function

As an alternative to approach 1, sismar also provides a wrapper function that calls both `clean_sisma_csv` and `parse_sisma_hiv_tarv` (or a similar user-specified parsing function). This reduces the amount of typing and code required for producing our final dataset. In this approach, we simply call `process_sisma_csv` and provide two required arguments: 1) The "file" argument that specifies the location of the SISMA .csv export and; 2) The "type" argument that specifies the programmatic data type for tidy'ing. A full list of values that can be passed to the "type" argument can be found in the `process_sisma_csv` [documentation](https://usaid-mozambique.github.io/sismar/reference/process_sisma_csv.html). 

```{r, eval = FALSE}
df <- process_sisma_csv(file = path_hiv_tarv, 
                        type = "HIV TARV")
```

Notice the output of this singular wrapper function is identical to the data frame generated in approach 1!
br<>
`process_sisma_csv` includes one more optional argument that defines the language of the variable names in the output file.  Currently the only language options are "portuguese" and "english" with the former defined as the argument default. By defining the argument value as "english" notice the change in the names of the variable output.

```{r, eval = FALSE}
df <- process_sisma_csv(file = path_hiv_tarv, 
                        type = "HIV TARV",
                        language = "english")

```

Happy analysis of your new tidy dataset!
